Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-06 demystify.tex
updates
using memoir class

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[margin=1.0in]{geometry}
\usepackage{soul}
%\usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=blue,
}
\usepackage{booktabs}

@@

\newenvironment{commentbox}[1][]{
\small
    \begin{mybox}
    {\small \textbf{#1}}
 }{
   \end{mybox}
}

\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}

\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}
%\date{\today}
\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\{\small George Mason University, Dept. of Computer Science}}

\begin{document}

\maketitle
\setcounter{section}{-1}

\begin{abstract}
    Having been involved in PhD admissions for many years, I've
    realized that many \textbf{international students}, especially those from smaller countries like Vietnam, lack a clear understanding of
    the Computer Science PhD admission process at US universities. This confusion not only
    discourages students from applying but also creates the perception that
    getting admitted to a CS PhD program in the US is difficult compared to other countries.

@@

    The latest version of document is at \url{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}. If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
\end{abstract}

\tableofcontents

\section{Summary}\label{sec:summary}

Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
\begin{enumerate}
\item Should you apply to a PhD in CS in the US?
\begin{itemize}
    \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
\end{itemize}
\item How is your application evaluated?
\begin{itemize}
    \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
    \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
\end{itemize}
\item Application Materials
\begin{itemize}
    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
    \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
    \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
\end{itemize}
\item What to do after getting admitted?
\begin{itemize}
    \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
\end{itemize}
\item Funding
\begin{itemize}
    \item TA and RA are two main sources of funding.  TA (teaching assistantship) is provided by the department to help profs. with classes (e.g., grading). RA (research assistantship) is provided by profs. to help them with their research (\S\ref{sec:funding}).
\end{itemize}
\item Choosing School and Professors
\begin{itemize}
    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be adviser, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
    \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
\end{itemize}
\item Miscs and FAQS
\begin{itemize}
    \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
    \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
    \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
    \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
    \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
\end{itemize}
\end{enumerate}

\section{Should You Apply to do a PhD in CS in the US?}\label{sec:should}

@@

\section{How is Your Application Evaluated?}\label{sec:evalapps}

@@

\section{Application}\label{sec:application}

@@

\subsection{Research Ability}

@@

\subsection{Letter of Recommendation (LOR)}\label{sec:lor}

@@

\subsection{Statement of Purpose (SOP)}\label{sec:research-statement}

@@

\subsection{Your School}\label{sec:your-school}

@@

\subsection{Grades/GREs}\label{sec:grades}

@@

\subsection{CV/Resume}
The CV/Resume should provide a summary of the applicant's achievements.  It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).

\subsection{Interview}
Sometimes a faculty wants to interview an applicant to make a decision. Typically, this means that they are considering admitting you. If they are not interested in your application, they will not proceed with an interview.

@@

\section{Getting Admitted}\label{sec:accepted}

@@

\section{Funding}\label{sec:funding}

@@

\subsection{Graduating Assistantship (TA/RA)}
The most common type of funding is \textbf{graduate assistanship}, which is either TA or RA. Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary). Some universities also pay insurance for spouse/children (or give very good discount).

@@

\subsubsection{Teaching Assistant (TA)}\label{sec:ta}

@@

\subsubsection{Research Assistant (RA)}
RA is provided through a professor through their own funding so you can work on their project.  
You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \S\ref{sec:ra-cost} gives more details on RA budget.

@@

\subsection{Fellowship/Scholarship}

@@

\section{Choosing Schools and Professors}\label{sec:schoolsandprofs}

@@

\subsection{Universities}

@@

\subsubsection{Schools that offer PhD in CS}

@@

%\subsubsection{R1, R2, ...}

\subsubsection{Selecting and Ranking Schools for CS}
   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

@@

\subsection{Choosing a Professor}

@@

\subsubsection{Finding the right adviser}
Obviously, there is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with, i.e., \emph{fits} you.  This is difficult and takes time.  Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.  This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find the right adviser.

\begin{commentbox}[Xiaokuan:]
    Whether the student's research interest matches that of the adviser is very important;
    if there is a mismatch,
    either the student or the adviser has to make compromises,
    which often leads to disagreements or conflicts. 
    IMO, the adviser should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
    instead of {\it dictating} their research.
    \end{commentbox}

@@

\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
Not every faculty in the CS program can serve as your formal adviser. Let's try to understand the different types of faculty and their roles. For example, you might hear about tenured,  tenure track, teaching faculty.  You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors. Here is a primer on these terminologies.

@@

\subsubsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}

@@

% \subsubsection{Strong Faculty vs High-ranked School}
% \begin{commentbox}[Thanh:]
% When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation? Of course, both are great, but it is hard to achieve both.  I believe receiving some guidance from you  would be incredibly valuable.
% \end{commentbox}

@@

\subsubsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}

@@

\begin{commentbox}[Xiaokuan]
    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic {\bf `why me, why you'} problem, 
similar to looking for a job in a company.
On a high level,
you need to show that you have done your {\it homework}
regarding the professor and the university,
and clearly explain:
1) why do you think you are a good fit in professor A's group?
2) why do you want to be advised by professor A, not B? 
3) why do you want to apply for university X, not Y?
If you don't want to spend time to do your homework,
the chance of getting a reply is close to zero.
\end{commentbox}

\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}

@@

\section{Miscs and FAQs}
\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}

\subsection{What can you do to increase your admission chance?}\label{sec:improve-your-chance}

@@

\subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}

@@

\subsection{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.

@@

\subsection{How long does it take to complete the CS PhD program?}\label{sec:time}

@@

\subsection{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.

@@

\subsection{How do I address a professor?}\label{sec:address}

@@

\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisers get. They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.

@@

%\subsection{Having fun during a PhD?}
%PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies. In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.

@@

\subsection{Will I be miserable during my a PhD?}\label{sec:happy}
Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable. Certainly issues including bad relationships with professors, conflicts with co-authors and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student) \emph{do} exist, and it is good to be aware of those.  However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.  So don't let social media make you feel pessimistic and deter your quest to advance knowledge.

\section{Cultural Differences and Other Issues}\label{sec:cultural}

@@

\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is quite different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

@@

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.  If you need software, ask your adviser or the department.

\paragraph{Gifts} In many countries, it is customary to give professors gifts, often during holidays.  These gifts can be costly and profs. sometimes expected them. In the US, it is uncommon and perfectly OK to not to give gifts. \emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated. Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences). In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.

@@

\section{Ranking of CS PhD programs in the U.S}\label{sec:ranking}
  Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.

  \begin{table}
    \centering
    \small
    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, June 2023). \red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
  \begin{tabular}{rl|rl}
    \toprule
    1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 &  Duke University \\
    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
    4 & MIT & 29 & Univ. of California - Riverside\\
    5 & Georgia Institute of Technology         & 30 & Northwestern University\\
    6 & Stanford University& 31 & Pennsylvania State University  \\
    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
    8 & University of Washington      &33 &  Harvard University \\
    9 &  Univ. of California - Berkeley  &34&  Univ. of California - Santa Cruz \\
    10 & Cornell University  & 35 &  Yale University \\
    11 & University of Maryland - College Park &  36& Brown University \\ 
    12 & Northeastern University\red{$^*$} &37&  Ohio State University\\
    13 & University of Wisconsin - Madison\red{$^*$}  &38& Texas A\&M University\red{$^*$} \\
    14 & Columbia University   &39 & Boston University  \\
    15 &   Purdue University  &40& North Carolina State University\\\  
    16 & University of Texas at Austin   &41 & University of Utah \\
    17 & University of Pennsylvania\red{$^*$} &42 & University at Buffalo\red{$^*$}\\
    18 & Princeton University  & 43& Rice University\\
    19 & University of Massachusetts-Amherst\red{$^*$} & 44&  University of Colorado-Boulder \\
    20 &  New York University  &45& University of Illinois at Chicago  \\
    21 & Univ. of California - Los Angeles &46& Virginia Tech\red{$^*$}  \\
    22 & University of Southern California &47&  Arizona State University\red{$^*$} \\
    23 & Stony Brook University\red{$^*$} &48&University of Minnesota \\
    24 & University of Chicago &49& University of Virginia \\
    25 & Univ. of California - Santa Barbara &50& University of North Carolina\red{$^*$} \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\section{History and Acknowledgement}

>>>>

\documentclass[oneside,11pt]{memoir}
\setsecnumdepth{subsection}

\usepackage[T1]{fontenc}
\usepackage[margin=1.2in]{geometry}
\usepackage{soul}
%\usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=blue,
}
\usepackage{booktabs}

@@

\newenvironment{commentbox}[1][]{
\small
    \begin{mybox}
    {\small \textbf{#1}}
 }{
   \end{mybox}
}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}

\title{Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}

\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\{\small George Mason University, Dept. of Computer Science}}

\begin{document}
\frontmatter

\maketitle
\chapter{Preface}
Having been involved in PhD admissions for many years, I've
realized that many \textbf{international students}, especially those from smaller countries like Vietnam, lack a clear understanding of
the Computer Science PhD admission process at US universities. This confusion not only
discourages students from applying but also creates the perception that
getting admitted to a CS PhD program in the US is difficult compared to other countries.

@@

The latest version of this document is at \url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}. If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.

\newpage
\tableofcontents*

\chapter{Summary}\label{sec:summary}
Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
\begin{enumerate}
\item Should you apply to a PhD in CS in the US?
\begin{itemize}
    \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
\end{itemize}
\item How is your application evaluated?
\begin{itemize}
    \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
    \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
\end{itemize}
\item Application Materials
\begin{itemize}
    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
    \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
    \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
\end{itemize}
\item What to do after getting admitted?
\begin{itemize}
    \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
\end{itemize}
\item Funding
\begin{itemize}
    \item TA and RA are two main sources of funding.  TA (teaching assistantship) is provided by the department to help profs. with classes (e.g., grading). RA (research assistantship) is provided by profs. to help them with their research (\S\ref{sec:funding}).
\end{itemize}
\item Choosing School and Professors
\begin{itemize}
    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be adviser, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
    \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
\end{itemize}
\item Miscs and FAQS
\begin{itemize}
    \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
    \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
    \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
    \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
    \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
\end{itemize}
\end{enumerate}

\mainmatter
\chapter{Should You Apply to a PhD program in CS in the US?}\label{sec:should}

@@

\chapter{How is Your Application Evaluated?}\label{sec:evalapps}

@@

\chapter{Application}\label{sec:application}

@@

\section{Research Ability}

@@

\section{Letter of Recommendation (LOR)}\label{sec:lor}

@@

\section{Statement of Purpose (SOP)}\label{sec:research-statement}

@@

\section{Your School}\label{sec:your-school}

@@

\section{Grades/GREs}\label{sec:grades}

@@

\section{CV/Resume}
The CV/Resume should provide a summary of the applicant's achievements.  It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).

\section{Interview}
Sometimes a faculty wants to interview an applicant to make a decision. Typically, this means that they are considering admitting you. If they are not interested in your application, they will not proceed with an interview.

@@

\chapter{Getting Admitted}\label{sec:accepted}

@@

\chapter{Funding}\label{sec:funding}

@@

\section{Graduating Assistantship (TA/RA)}
The most common type of funding is \textbf{graduate assistanship}, which is either TA or RA. Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary). Some universities also pay insurance for spouse/children (or give very good discount).

@@

\subsection{Teaching Assistant (TA)}\label{sec:ta}

@@

\subsection{Research Assistant (RA)}
RA is provided through a professor through their own funding so you can work on their project.  
You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \S\ref{sec:ra-cost} gives more details on RA budget.

@@

\section{Fellowship/Scholarship}

@@

\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}

@@

\section{Choosing a University}

@@

\subsection{Schools that offer PhD in CS}

@@

%\subsection{R1, R2, ...}

\subsection{Selecting and Ranking Schools for CS}
   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

@@

\section{Choosing a Professor}

@@

\subsection{Finding the right adviser}
Obviously, there is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with, i.e., \emph{fits} you.  This is difficult and takes time.  Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.  This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find adviser.

\begin{commentbox}[Xiaokuan:]
    Whether the student's research interest matches that of the adviser is very important;
    if there is a mismatch,
    either the student or the adviser has to make compromises,
    which often leads to disagreements or conflicts. 
    IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
    instead of \emph{dictating} their research.
    \end{commentbox}

@@

\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
Not every faculty in the CS program can serve as your formal adviser. Let's try to understand the different types of faculty and their roles. For example, you might hear about tenured,  tenure track, teaching faculty.  You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors. Here is a primer on these terminologies.

@@

\subsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}

@@

% \subsection{Strong Faculty vs High-ranked School}
% \begin{commentbox}[Thanh:]
% When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation? Of course, both are great, but it is hard to achieve both.  I believe receiving some guidance from you  would be incredibly valuable.
% \end{commentbox}

@@

\subsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}

@@

\begin{commentbox}[Xiaokuan]
    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic \textbf{`why me, why you'} problem, 
similar to looking for a job in a company.
On a high level,
you need to show that you have done your \emph{homework}
regarding the professor and the university,
and clearly explain:
1) why do you think you are a good fit in professor A's group?
2) why do you want to be advised by professor A, not B? 
3) why do you want to apply for university X, not Y?
If you don't want to spend time to do your homework,
the chance of getting a reply is close to zero.
\end{commentbox}

\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}

@@

\chapter{Miscs and FAQs}
\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}

\section{What can you do to increase your admission chance?}\label{sec:improve-your-chance}

@@

\section{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}

@@

\section{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.

@@

\section{How long does it take to complete the CS PhD program?}\label{sec:time}

@@

\section{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.

@@

\section{How do I address a professor?}\label{sec:address}

@@

\section{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisers get. They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.

@@

%\section{Having fun during a PhD?}
%PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies. In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.

@@

\section{Will I be miserable during my a PhD?}\label{sec:happy}
Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable. Certainly issues including bad relationships with professors, conflicts with co-authors and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student) \emph{do} exist, and it is good to be aware of those.  However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.  So don't let social media make you feel pessimistic and deter your quest to advance knowledge.

\chapter{Cultural Differences and Other Issues}\label{sec:cultural}

@@

\section{Academic Integrity}

Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is quite different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

@@

\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.  If you need software, ask your adviser or the department.

\section{Gifts} In many countries, it is customary to give professors gifts, often during holidays.  These gifts can be costly and profs. sometimes expected them. In the US, it is uncommon and perfectly OK to not to give gifts. \emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated. Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences). In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.

@@

\appendix

\chapter{Ranking of CS PhD programs}\label{sec:ranking}

Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
    \begin{table}
    \centering
    \small
    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, June 2023). \red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
  \begin{tabular}{rl|rl}
    \toprule
    1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 &  Duke University \\
    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
    4 & MIT & 29 & Univ. of California - Riverside\\
    5 & Georgia Institute of Technology         & 30 & Northwestern University\\
    6 & Stanford University& 31 & Pennsylvania State University  \\
    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
    8 & University of Washington      &33 &  Harvard University \\
    9 &  Univ. of California - Berkeley  &34&  Univ. of California - Santa Cruz \\
    10 & Cornell University  & 35 &  Yale University \\
    11 & University of Maryland - College Park &  36& Brown University \\ 
    12 & Northeastern University\red{$^*$} &37&  Ohio State University\\
    13 & University of Wisconsin - Madison\red{$^*$}  &38& Texas A\&M University\red{$^*$} \\
    14 & Columbia University   &39 & Boston University  \\
    15 &   Purdue University  &40& North Carolina State University\\\  
    16 & University of Texas at Austin   &41 & University of Utah \\
    17 & University of Pennsylvania\red{$^*$} &42 & University at Buffalo\red{$^*$}\\
    18 & Princeton University  & 43& Rice University\\
    19 & University of Massachusetts-Amherst\red{$^*$} & 44&  University of Colorado-Boulder \\
    20 &  New York University  &45& University of Illinois at Chicago  \\
    21 & Univ. of California - Los Angeles &46& Virginia Tech\red{$^*$}  \\
    22 & University of Southern California &47&  Arizona State University\red{$^*$} \\
    23 & Stony Brook University\red{$^*$} &48&University of Minnesota \\
    24 & University of Chicago &49& University of Virginia \\
    25 & Univ. of California - Santa Barbara &50& University of North Carolina\red{$^*$} \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\chapter{History and Acknowledgement}
